---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pipe

This package proposes an alternative to the pipe from 'magrittr'. 
It's named the same and passes all the 'magrittr' test so can easily be a drop-in
replacement, its main advantages is that it is implemented more consistently and allows definition of new pipes. It also solves a few issues of the original pipe. 

Install with :

``` r
remotes::install_github("pipe")
```
## What's different from *magrittr* ?

The main differences are:

* Pipes are object of class `pipe` and of class either `standard_pipe` or `compound_pipe`, none of them is special implementation-wise

* Every step of the chain is executed in the same environment, as can be seen
 by the use of `%B>%` above
 
* The new pipe `%B>%` allows browsing the pipe chain step by step

calling `iris %>% head %B>% dim %>% length` will open place you right at the browser call below :

```{r, eval = FALSE}
#> Called from: (function (.) 
#> {
#>     on.exit(rm(.))
#>     . <- head(.)
#>     browser()
#>     . <- dim(.)
#>     . <- length(.)
#>     .
#> })(iris)
```

* The new pipe `%S>%` allows the use of *rlang*'s `!!!` operator to splice dots in
any function 

```{r}
library(pipe)
c(a = 1, b = 2) %S>% data.frame()
```

* additional pipes can be defined easily by calling `new_pipe()`, which is used
internally to create pipes included in the package

* `:::`, `::` and `$` get a special treatment to solve a common issue

```{r}
iris %>% base::dim
iris %>% base:::dim
x <- list(y = dim)
iris %>% x$y
```

* It fails explicitly in some cases rather than allowing strange behavior silently

```{r, error = TRUE}
iris %>% head %<>% dim
. %<>% head
```


* It's slightly slower than *magrittr* (see *fast_pipe* if speed is an issue!)

```{r}
microbenchmark::microbenchmark(
  magrittr = {`%>%` <- magrittr::`%>%`
  1 %>% identity %>% identity() %>% (identity) %>% {identity(.)}},
  pipe = {`%>%` <- pipe::`%>%`
  1 %>% identity %>% identity() %>% (identity) %>% {identity(.)}},
  times=10000
)
```

## Defining new pipes

Pipes are defined by the function `new_pipe()` which accepts arguments `format_fun` and `compound_fun`.

`format_fun` is the function that changes the quoted rhs of a pipe into the quoted expression that will give a new value to the dot. For instance in `iris %>% head()` its input will be `quote(head())` and it's output `quote(. <- head(.))`.

`compound_fun` is used to define new types of compound pipes, we won't detail
it as it's very specific, but see last example, used to define `%<>%`.

We enumerate all the pipes of the package below :

```{r}
`%>%` <- new_pipe(
  function(call){
  call <- insert_dot(call)
  bquote(. <- .(call))
})

`%T>%`<- new_pipe(
  function(call){
  call <- insert_dot(call)
  bquote(local(.(call)))
})

`%$%`<- new_pipe(
  function(call){
  bquote(. <- with(.,.(call)))
})

`%S>%`<- new_pipe(
  function(call){
    call <- insert_dot(call)
    bquote(. <- eval(rlang::expr(.(call))))
  })


`%B>%` <- new_pipe(
  function(call){
  call <- insert_dot(call)
  c(quote(browser()),
    bquote(. <- .(call)))
})

`%<>%` <- new_pipe(
  function(call){
    call <- insert_dot(call)
    bquote(. <- .(call))
  },
  function(lhs, res){
    substitute(lhs <- res)
  })
```

## A few notes 

This is just an experiment, it's likely to change. I've also been working on the
package *fast_pipe* which is a closer equivalent to *magrittr* focused on
being more robust and faster.

I've worked on two pipe packages in the past, names *mmpipe* (archived) and *pipes*
which were forks of `magrittr`. 

I believe the current approach is stronger and that the package *pipes* should use *pipe* to define its operators and that `%B>%` and `%S>%` so that the present package contain only the equivalents original *magrittr* pipes.
